Cultural Evolution of Conservation Behavior at the Intersection of Working

### 2020-09-30 DissertationProposal/DissertationProposal.Rmd
finished cites for now

# Introduction
## Phenomenon
Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources. Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources (GFN, 2019). Today, on average, we degenerate these resources at a rate 56% faster than regeneration (Lin et al., 2020). This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining (Beltratti et al., 1993). Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity (Regev et al., 1998).

The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity (Haines-Young & Potschin, 2010, Redford, 1995). As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target (Parrish et al., 2003). Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) (UNEP-WCMC & IUCN 2020). These area-based conservation initiatives are from hereon referred to as protected areas.

@@

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations (@adams_biodiversity_2004) (Leverington et al., 2010). While high-profile protected area successes are salient in the conservation narrative, these areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported (Bottrill et al., 2011; Sutherland et al., 2004). Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations (Catalano et al., 2019).

Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group (Rogers, 2010; Simon et al., 2012). These behaviors are characterized as individually costly, yet beneficial at the group level. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success (Bennett et al., 2017). This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present (Ostrom, 1990; Richerson & Boyd, 2001). What is less understood however, is the emergence and initial adoption of conservation behaviors.

@@

## Primer on cultural group selection
The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations (Boyd & Richerson, 2010). Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not. As an individual’s behavior is flexible, individuals will either migrate from less to more successful groups or directly copy the cooperative behavior of others in visibly successful cooperative groups (Henrich, 2004). This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection. In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall (McElreath & Boyd, 2008; Okasha, 2004; Price, 1972).

Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors (Boyd et al., 2011; Lehmann et al., 2008). Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically. Conservation behaviors are fundamentally different from generic cooperative behaviors as they operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates. As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.

Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals (Adams et al., 2004). While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice (Naughton-Treves et al., 2005). Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

## Fidelity of behavioral transmission
The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission. In situations where behaviors do emerge and spread within a population, as the behavior is transmitted along the chain of individuals, it may be modified considerably to the point where the behavioral outcome is unrecognizable (Lewis & Laland, 2012; McGuigan et al., 2011). This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves. Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator (King & Cowlishaw, 2007; Muthukrishna et al., 2016). These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.

@@

# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
## Background
Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) (UNEP-WCMC & IUCN 2020). During this time period, the stated mission of most protected areas has shifted as well. Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities (Fisher & Christopher, 2007; Naughton-Treves et al., 2005). With this recognition, their objectives now commonly include community development goals alongside conservation targets.

@@

Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss (Tallis et al., 2008). These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations (Hall & Tucker, 2004; Langholz & Lassoie, 2001). Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not (Fisher & Christopher, 2007). Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation (Adams et al., 2004).

Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals (Christie, 2004; Watson et al., 2014). In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands (Clark et al., 2013). It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations (Bennett & Dearden, 2014; Christie, 2005).

The degree of local buy-in for protected areas is not random. The scale of implementation is highly correlated with the percentage of individuals who adopt protected area regulations. A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters (Mills et al., 2019). These findings are particularly worrying given a global decline in government support for land conservation (Watson et al., 2014).

With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research (Bennett et al., 2017). This body of work has largely relied on case studies of conservation projects to draw correlations between social phenomena and conservation outcomes. As such, findings from conservation social science tend to offer little mechanistic insights for how conservation behaviors spread within and between groups. Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives (Brooks et al., 2018).

Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups (Rindos et al., 1985). Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge (Richerson et al., 2016).

@@

## Background
Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection (Molleman et al., 2013). One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use (Waring et al., 2017). Under this prediction, selection acts both on individuals and on groups of individuals. When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection. Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups (Boyd & Richerson, 2010; Henrich, 2004).

While these predictions are theoretically sound, they must be tested empirically. Two such studies have begun to test the basis of cultural group selection in the natural world. The first demonstrated that culture offers greater fodder for group selection than do genetics. Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals. The proportion of variance in an allele or trait found between groups is referred to as genetic Fst or cultural Fst respectively. Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits (Bell et al., 2009).

A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups. Researchers created pair-wise estimates of cultural Fst between 12 ethnic groups. These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups. Results of this study showed a strong association between cultural similarity and willingness to cooperate. This research indicates that the presence of variable cultural groups influences within-group cooperative behavior (Handley & Mathew, 2020).

@@

## Expected Outputs
The primary outputs of data collection and analysis in this project will produce binary predictions of protected area adoption for each study ward, as well as the calculated strength of selection for protected area adoption for each study ward. I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of community managed protected areas. (Figure \@ref(fig:aelPress)) shows an example of what the results of this study might look like. The findings of this study will be submitted for publication in a peer reviewed, academic journal.

@@

## Background
Forests sequester 2 billion tons of atmospheric CO2 each year, while also providing fuel to 40% of the world’s rural poor — fuelwood produces more renewable energy than solar, wind, and hydroelectric sources combined (SOFO, 2018). Household fuel production and global carbon sequestration are often seen as conflicting forest benefits. Solutions to mitigate this tradeoff include expanding markets for non-timber forest products and technological advances like solar panels (McShane et al., 2011). Given limited funding, these projects are not intended to solve problems in their entirety, but rather to seed innovations that catalyze broad change. Many of these initiatives, however, fail to spread and produce long-term benefits for local livelihoods or forest conservation because they do not account for underlying social dynamics (Keppel et al., 2012; Oates, 1999; Stringer & Reed, 2007).

This project will examine the spread of fuel-efficient cookstoves on Pemba Island, Tanzania (see Figure \@ref(fig:stovePic)). An improved stove design which increases fuel efficiency by up to 50%, was seeded in three communities on the island in 2015 (Lung & Espira, 2019; Sakthivadivel & Iniyan, 2018). The improved design has been subsequently copied throughout the communities and is thought to reduce deforestation and the time individuals — primarily women and children — must spend harvesting fuelwood.

@@

While this project will inform effective implementation of household level conservation projects on the ground, it will also provide important insights for understanding behavioral adoption more generally. While the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially (Lyons et al., 2007; McGuigan et al., 2011). In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly (Bartlett, 1932). Behavioral reconstruction creates variation in copied behaviors (Horner et al., 2006). Replication variation is thought to be greatly influenced by environmental factors, including social context (Mesoudi & Whiten, 2008).

Both theoretical and experimental studies have examined the relationship between transmission fidelity and environmental influences. Theoretical work suggests that increasing transmission fidelity should be associated with greater reliance on social learning (Boyd & Henrich, 2002; Nakahashi et al., 2012). This theory has also been validated experimentally (Muthukrishna et al., 2016). When coupled with the theoretical prediction that social learning should increase with greater access to behavioral demonstrators, it follows logically that transmission fidelity should increase with imitator access to the behavioral model (King & Cowlishaw, 2007).

While transmission fidelity can vary significantly given the context, it is often unclear what effect this variation has on the outcomes of imitated behavior. Theoretical models have shown that even small changes in transmission fidelity can have strong impacts on population-level behavioral outcomes (Lewis & Laland, 2012). Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes (Enquist et al., 2010). In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process.

Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity. Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes (Mesoudi & Whiten, 2008). This research will directly measure the impact of geographic, genetic, and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.

@@

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig XXX). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

@@

While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics such as the exterior dimensions and size of the openings. I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from. To measure stove fuel efficiency, I will follow the standard water boiling test protocol (DeFoort et al., 2009; McCracken & Smith, 1998). Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test.

@@

## Preliminary Work
In preparation for a cancelled field season in the Summer of 2020, I created a variety of forms and documents to facilitate data collection. These include:
* A household livelihood and stove adoption survey.  
* A datasheet for cookstove fuel efficiency measurements.
* A fuel efficiency measurement protocol. 
I also simulated data that matches the data would be collected by the multiple instruments. These data were used to test the ability of my proposed statistical methods to recover the signal of demonstrator access on behavioral transmission fidelity and behavioral outcomes.

>>>>

# Introduction
## Phenomenon
Human prosperity is unequivocally linked to the persistence of diverse, self-sustaining ecosystems through our dependence on ecosystem services and consumption of renewable resources. Until approximately 50 years ago, the rate of overall ecological regeneration exceeded our rates of consumption and destruction of earth’s renewable resources [@network2019global]. Today, on average, we degenerate these resources at a rate 56% faster than regeneration [@network2020global]. This is particularly concerning as renewable resources regenerate at a rate proportional to the resource stock remaining [@beltratti_sustainable_1993]. Thus, overuse of earth’s renewable resources creates a positive feedback loop where scarcity slows regeneration and further perpetuates scarcity [@regev_biological_1998].

The capacity of an ecosystem to provision renewable resources largely hinges on an area’s biodiversity [@haines2010links; @redford1995human]. As such, biodiversity is the key metric used to determine the integrity of an ecosystem and a key conservation target [@parrish_are_2003]. Each year, an ever growing proportion of earth’s marine and terrestrial areas are sectioned off to protect local biodiversity (hotspots), the flow of genetic diversity across a landscape (corridors), and human access to renewable resources (forest concessions, fisheries, etc.) (Figure \@ref(fig:paNumber)) [@iucn2020protectedplanet]. These area-based conservation initiatives are from hereon referred to as protected areas.

@@

Protected areas have proven to be a powerful tool for sustaining biodiversity and ecosystem function at a landscape scale while also ensuring continued access to renewable resources for local human populations [@adams_biodiversity_2004; @leverington_global_2010]. While high-profile protected area successes are salient in the conservation narrative, these areas also commonly fail to reach their intended biodiversity targets. The ubiquity of these failures is highly speculative as unsuccessful protected area initiatives go largely unreported [@bottrill_pursuit_2011; @sutherland_need_2004]. Despite a high degree of uncertainty in the prevalence of protected area failures, the cause is more often than not a lack of social buy-in and adherence to protected area regulations [@catalano_learning_2019].

Community adoption of protected area regulations may emerge and persist as a group norm, emerge only for a short period of time, or may never emerge within a given group [@rogers_diffusion_2010; @simon_win-win_2012]. These behaviors are characterized as individually costly, yet beneficial at the group level. Given that the human dimensions are most often the crux of successful conservation, the field of conservation social science has been developed to better understand which social factors are pertinent to conservation success [@bennett_conservation_2017]. This body of work has shown that where group norms around the adoption of protected areas have emerged, they persist via group conformity, especially where clear group identities are present [@ostrom_governing_1990; @richerson2001evolution]. What is less understood however, is the emergence and initial adoption of conservation behaviors.

@@

## Primer on cultural group selection
The largely untested theory of cultural group selection poses a mechanism for the emergence and spread of individually costly yet group beneficial behaviors such as warfare or religious donations [@boyd_transmission_2010]. Cultural group selection suggests that groups who develop individually costly yet group beneficial norms will outcompete groups who do not. As an individual’s behavior is flexible, individuals will either migrate from less to more successful groups or directly copy the cooperative behavior of others in visibly successful cooperative groups [@henrich_cultural_2004]. This change in the frequency of behavioral traits is analogous to the perpetuation of beneficial alleles in genetic natural selection. In these contexts, selective pressures still push individuals towards selfish behaviors like freeriding, but cooperative norms will emerge if group selection pressures outweigh those of individual selection overall [@mcelreath_mathematical_2008; @okasha_multilevel_2004; @price1972extension].

Existing cultural group selection theory is highly abstract, relying on simple, analytical models of nonspecific behaviors [@boyd_rapid_2011; @lehmann_cultural_2008]. Current theory therefore does not offer predictions about the emergence and adoption of conservation norms specifically. Conservation behaviors are fundamentally different from generic cooperative behaviors as they operate under the sink/source dynamics of renewable, mobile resources as well as variable resource regeneration rates. As such, cultural group selection theory must be extended to include these dynamics in order to be effectively applied to conservation.

Applying cultural group selection to the adoption of conservation behaviors
Protected areas are an ideal microcosm for studying and applying cultural group selection. Protected areas have the potential to support resource sustainability and local livelihoods at the group level, but require costly cooperation from individuals [@adams_biodiversity_2004]. While this tradeoff is straightforward in theory, the social and ecological outcomes of protected area establishment vary significantly in practice [@naughton-treves_role_2005]. Agrawal and Redford [-@agrawal2006poverty] argue that this is because protected areas are often created as “a shot in the dark,” with little attention paid to the mechanisms which cause protected areas to be widely adopted and subsequently successful. Under cultural group selection theory, we expect that the key mechanism which drives the adoption of protected areas is the balance between the individual and group level selection pressures for cooperation or defection. Therefore, not only do protected areas represent a system where cultural group selection theory can be studied empirically, but principles of cultural group selection may also provide a way forward for predicting and shaping the success of protected areas on the ground.

## Fidelity of behavioral transmission
The final component of behavioral adoption that is necessary to understand and shape effective conservation is the fidelity of behavioral transmission. In situations where behaviors do emerge and spread within a population, as the behavior is transmitted along the chain of individuals, it may be modified considerably to the point where the behavioral outcome is unrecognizable [@lewis_transmission_2012; @mcguigan_over-imitation_2011]. This phenomenon is of particular interest to conservation science where the outcomes of behavioral adoption are of greater interest than the behaviors themselves. Theoretical insights and controlled empirical research on transmission fidelity suggests that behavioral outcomes are more likely to be retained during transmission when the copying individual has greater access to the demonstrator [@king2007use; @muthukrishna_when_2016]. These insights are crucial for practitioners aiming to implement long-lasting, effective conservation projects, but must first be tested empirically in a conservation context where an individual’s access to a behavioral demonstration is realistic and outcomes of adopted behaviors have real impacts for individuals, communities, and the environment.

@@

# Chapter 1: Theoretical Advances for Applying Cultural Group Selection Theory to Area-Based Conservation. 
## Background
Over the last 20-30 years, both the number and size of new no-take protected areas created each year have increased steadily (Figures \@ref(fig:paNumber) & \@ref(fig:paSize)) [@iucn2020protectedplanet]. During this time period, the stated mission of most protected areas has shifted as well. Due to the overlap between biodiversity and low income populations, as well as the price and availability of land, protected areas are largely adjacent to, or even contain, low income communities [@fisher_poverty_2007; @naughton-treves_role_2005]. With this recognition, their objectives now commonly include community development goals alongside conservation targets.

@@

Joint conservation and community development initiatives are a popular prescription for combating climate change and global biodiversity loss [@tallis_ecosystem_2008]. These dual-objective projects attempt to mitigate concerns of environmental neocolonialism and resource disenfranchisement of local populations [@hall_tourism_2004; @langholz_perils_2001]. Importantly, protected areas which explicitly seek to bolster the livelihoods of local populations actually show greater conservation success than those that do not [@fisher_poverty_2007]. Protected areas then represent a considerable opportunity for “win-win” projects which support global conservation and poverty alleviation [@adams_biodiversity_2004].

Protected areas, while well intentioned, often fail to meet both their stated social and ecological goals [@christie_marine_2004; @watson_performance_2014]. In fact, a study of protected areas in South Asia found them to be ecologically equivalent to adjacent unprotected lands [@clark_protected_2013]. It’s unsurprising that the success of protected areas in reaching both social and ecological targets has been linked to the degree of local support and adherence to regulations [@bennett_why_2014; @christie_observed_2005].

The degree of local buy-in for protected areas is not random. The scale of implementation is highly correlated with the percentage of individuals who adopt protected area regulations. A 2019 study demonstrated that nationally implemented initiatives had 84-100% buy-in from local populations, while locally established areas saw adoption from 0.13 - 51% of potential adopters [@mills_how_2019]. These findings are particularly worrying given a global decline in government support for land conservation [@watson_performance_2014].

With the recognition that local adoption is a difficult and necessary component of conservation success, human dimensions have become a major focus of conservation research [@bennett_conservation_2017]. This body of work has largely relied on case studies of conservation projects to draw correlations between social phenomena and conservation outcomes. As such, findings from conservation social science tend to offer little mechanistic insights for how conservation behaviors spread within and between groups. Recently however, theories from cultural group selection have been identified as a potentially useful tool to better elucidate the mechanisms which underpin local buy-in for conservation initiatives [@brooks_applying_2018].

Cultural group selection theorizes that individuals will adopt behaviors which adhere to protected area regulations if the group benefits of doing so allow their cultural group to outcompete other groups [@rindos_darwinian_1985]. Under this framework, multiple competing groups are actually required for community adoption of protected area regulations to emerge [@richerson_cultural_2016].

@@

## Background
Cultural group selection offers specific predictions about when individually costly, yet group beneficial norms will be favored by selection [@molleman_cultural_2013]. One such prediction is that the presence of multiple competing groups is the key to the development of norms for sustainable resource use [@waring_coevolution_2017]. Under this prediction, selection acts both on individuals and on groups of individuals. When the ratio of group benefit to individual cost of cooperation outweighs the individual fitness consequences of cooperation, those behaviors will be favored by selection. Groups who develop cooperative norms around natural resource use will outcompete other groups, who may then copy the cooperative institutions established by the more successful groups [@boyd_transmission_2010; @henrich_cultural_2004].

While these predictions are theoretically sound, they must be tested empirically. Two such studies have begun to test the basis of cultural group selection in the natural world. The first demonstrated that culture offers greater fodder for group selection than do genetics. Researchers estimated the total variance in both cultural and genetic traits within and between groups of individuals. The proportion of variance in an allele or trait found between groups is referred to as genetic Fst or cultural Fst respectively. Findings indicate that between human groups, there is greater differentiation in culture than genes, indicating that group selection will act more readily on cultural traits [@bell_culture_2009].

A second empirical study tested the prediction that large scale human cooperation is in fact a product of competition between cultural groups. Researchers created pair-wise estimates of cultural Fst between 12 ethnic groups. These estimates of cultural differentiation were combined with a survey of individual willingness to cooperate with members of the other groups. Results of this study showed a strong association between cultural similarity and willingness to cooperate. This research indicates that the presence of variable cultural groups influences within-group cooperative behavior [@handley_human_2020].

@@

## Expected Outputs
The primary outputs of data collection and analysis in this project will produce binary predictions of protected area adoption for each study ward, as well as the calculated strength of selection for protected area adoption for each study ward. I will compare these calculated adoption likelihoods and selection strengths against actual adoption rates of community managed protected areas. (Figure \@ref(fig:selPress)) shows an example of what the results of this study might look like. The findings of this study will be submitted for publication in a peer reviewed, academic journal.

@@

## Background
Forests sequester 2 billion tons of atmospheric CO2 each year, while also providing fuel to 40% of the world’s rural poor — fuelwood produces more renewable energy than solar, wind, and hydroelectric sources combined [@muller2018state]. Household fuel production and global carbon sequestration are often seen as conflicting forest benefits. Solutions to mitigate this tradeoff include expanding markets for non-timber forest products and technological advances like solar panels [@mcshane_hard_2011]. Given limited funding, these projects are not intended to solve problems in their entirety, but rather to seed innovations that catalyze broad change. Many of these initiatives, however, fail to spread and produce long-term benefits for local livelihoods or forest conservation because they do not account for underlying social dynamics [@keppel_conservation_2012; @oates_myth_1999; @stringer_land_2007].

This project will examine the spread of fuel-efficient cookstoves on Pemba Island, Tanzania (see Figure \@ref(fig:stovePic)). An improved stove design which increases fuel efficiency by up to 50%, was seeded in three communities on the island in 2015 [@lung_large-scale_2019; @sakthivadivel2018experimental]. The improved design has been subsequently copied throughout the communities and is thought to reduce deforestation and the time individuals — primarily women and children — must spend harvesting fuelwood.

@@

While this project will inform effective implementation of household level conservation projects on the ground, it will also provide important insights for understanding behavioral adoption more generally. While the ubiquity and utility of behavioral imitation, or copying, is well documented, the fidelity of behavioral imitation varies circumstantially [@lyons_hidden_2007; @mcguigan_over-imitation_2011]. In a foundational study, researchers show that learned behavior is reconstructed and assimilated with an individual’s prior experiences, rather than replicated directly [@bartlett1932remembering]. Behavioral reconstruction creates variation in copied behaviors [@horner_faithful_2006]. Replication variation is thought to be greatly influenced by environmental factors, including social context [@mesoudi_multiple_2008].

Both theoretical and experimental studies have examined the relationship between transmission fidelity and environmental influences. Theoretical work suggests that increasing transmission fidelity should be associated with greater reliance on social learning [@boyd_modeling_2002; @nakahashi_adaptive_2012]. This theory has also been validated experimentally [@muthukrishna_when_2016]. When coupled with the theoretical prediction that social learning should increase with greater access to behavioral demonstrators, it follows logically that transmission fidelity should increase with imitator access to the behavioral model [@king_when_2007].

While transmission fidelity can vary significantly given the context, it is often unclear what effect this variation has on the outcomes of imitated behavior. Theoretical models have shown that even small changes in transmission fidelity can have strong impacts on population-level behavioral outcomes [@lewis_transmission_2012]. Alternatively, some behaviors require only very low-fidelity replication to achieve identical outcomes [@enquist_one_2010]. In fact, it is likely that if a behavior is under strong environmental constraints then individual learning should help correct errors in the transmission process.

Currently, we do not yet understand how various forms of access to behavioral demonstration may affect transmission fidelity. Secondly, relatively little empirical research has examined the relationship between transmission fidelity and behavioral outcomes [@mesoudi_multiple_2008]. This research will directly measure the impact of geographic, genetic, and social relatedness on the fidelity of copying behavior, and the impact of transmission fidelity on specific behavioral outcomes.

@@

### Methodological Specifics
I will map all the seeded and voluntarily copied fuel-efficient cookstoves in Makangale, Msituu mkuu, and Kifundi wards on Pemba Island, Tanzania (Fig \@ref(fig:mapp)). With the assistance of a translator, I will survey all the individuals responsible for constructing each cookstove, measure a suite of physical characteristics of each cookstove (length, height, etc.), and measure the fuel efficiency.

@@

While conducting the fuel efficiency test, I will also measure a variety of physical stove characteristics such as the exterior dimensions and size of the openings. I will use these measures to produce quantitative estimates of how physically similar each cookstove is to the cookstove it was copied from. To measure stove fuel efficiency, I will follow the standard water boiling test protocol [@defoort2009stove; @mccracken1998emissions]. Water temperature will be measured in standard intervals and compared against the weight of fuel-wood before and the weight of fuel-wood and charcoal remaining at the end of the test.

@@

## Preliminary Work
In preparation for a cancelled field season in the Summer of 2020, I created a variety of forms and documents to facilitate data collection. These include:

* A household livelihood and stove adoption survey.

* A datasheet for cookstove fuel efficiency measurements.

* A fuel efficiency measurement protocol.

I also simulated data that matches the data would be collected by the multiple instruments. These data were used to test the ability of my proposed statistical methods to recover the signal of demonstrator access on behavioral transmission fidelity and behavioral outcomes.
